### 2018-05-04 template.Rmd
rmd template with knitr settings

---
title: 'Title'
author: '[Ying Li](https://business.illinois.edu/profile/ying-li/)'
date: '`r format(Sys.Date())`'
fontsize: 12pt
geometry: margin=1in
output:
  pdf_document:
    latex_engine: xelatex
    fig_caption: yes
    fig_height: 4
    fig_width: 4
---

```{r globalsettings, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  tidy=TRUE,
  echo=FALSE,
  message=FALSE,
  results='asis',
  strip.white=TRUE,
  fig.path='Figures/',
  cache=FALSE,
  highlight=TRUE,
  width.cutoff=80,
  size='footnotesize',
  out.width='.9\\textwidth',
  fig.retina=FALSE,
  comment=NA,
  out.extra="")

if(!file.exists('Figures/')) dir.create('Figures/')
```

```{r eval=FALSE}
# Frequentely used:
- Insert a figure
\begin{figure}
\includegraphics{bayesplot.png}
\caption{Estimates and Credible Intervals of A Bayesian Logit Model}
\label{fig:bayesplot}
\end{figure}
Figure \ref{fig:bayesplot}
```