### 2018-05-15 template.Rmd

---
title: 'Title'
author: '[Ying Li](https://business.illinois.edu/profile/ying-li/)'
date: '`r format(Sys.Date(), "%B %d, %Y")`'
fontsize: 12pt
geometry: margin=1in
header-includes: \usepackage{setspace}\doublespacing
output:
  pdf_document:
    latex_engine: xelatex
    fig_caption: yes
    fig_height: 4
    fig_width: 4
---

```{r globalsettings, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  tidy=TRUE,
  tidy.opts=list(width.cutoff=80),
  echo=FALSE,
  message=FALSE,
  warning=FALSE,
  error=FALSE,
  collapse = TRUE,
  highlight=TRUE, # Increase the readability of source code syntax
  results='asis',
  strip.white=TRUE, #To skip white lines
  fig.path='Figures/',
  cache=TRUE,
  cache.path='cache/',
  award_time=file.info('award.csv')$mtime, # data will be re-read if award.csv becomes newer
  width.cutoff=80,
  size='footnotesize', # Other options include small, large, Large
  out.width='.9\\textwidth',
  fig.retina=FALSE,
  comment=NA,
  out.extra="")
if(!file.exists('Figures/')) dir.create('Figures/')
options(scipen = 1, digits = 2)
```

```{r eval=FALSE}
# Frequentely used:
- Insert a figure
\begin{figure}
\includegraphics{bayesplot.png}
\caption{Estimates and Credible Intervals of A Bayesian Logit Model}
\label{fig:bayesplot}
\end{figure}
Figure \ref{fig:bayesplot}

# Appendix 
```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
If you want to remove the first chunk, you can write knitr::all_labels()[-1].
Of course, this requires you to give a label to each chunk.
```

# Appendix^[I ]
```{r appendix, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
```